---
title: "The Effects of Vitamin C Delivery Methods on Tooth Growth"
author: "Steve Pederson"
date: "Wednesday, August 19, 2015"
output: word_document
---

# Required Packages

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
```

# Data Description

We are interested in the effects of Vitamin C on tooth growth, and for this experiment we used Guinea Pigs as the model organism.
We gave `r nrow(ToothGrowth)` guinea pigs Vitamin C supplements at three doses (0.5, 1 and 2*mg*) using two delivery methods.
Delivery method 1 was orange juice (OJ), whilst delivery method 2 was ascorbic acid (VC).
10 guinea pigs were given each combinations of treatments, and the measured response was the length of odontoblasts.

```{r}
ToothGrowth %>% 
  group_by(supp, dose) %>% 
  summarise(Count = n()) 
```

```{r, echo =FALSE, eval=FALSE}
ggplot(ToothGrowth, aes(x = as.factor(dose), y = len)) +
  geom_boxplot() +
  facet_grid(~supp) + 
  labs(x = "Dose (mg)", y = "Odontoblast Length (mm)")
```

```{r, echo=FALSE}
ggplot(ToothGrowth, aes(x = log2(dose), y = len, colour = supp)) +
  geom_point() +
  facet_grid(~supp) + 
  labs(x = "Dose (mg)", y = "Odontoblast Length (mm)")
```

# Analysis

To analyse this dataset, we conducted $t$-tests comparing each delivery method at each dosage level, testing the null hypothesis <br>

$H_0$: there is no effect on odontoblast length as a result of Vitamin C delivery method

with the alternate hypothesis

$H_A$: there is an effect on odontoblast length as a result of Vitamin C delivery method

## Dosage  = 0.5
```{r}
(lowDosage <- t.test(len~supp, subset(ToothGrowth, dose==0.5)))
```

## Dosage  = 1
```{r}
(medDosage <- t.test(len~supp, subset(ToothGrowth, dose==1)))
```

## Dosage  = 2
```{r}
(highDosage <-t.test(len~supp, subset(ToothGrowth, dose==2)))
```

# Results

## Significance tests
As seen in the final of the $t$-tests above, there is no significant difference between delivery methods at the highest dosage level (dose==2; $p=$ `r round(highDosage$p.value, 4)`).
However, significant differences were observed at both the moderate level (dose==1; $p=$ `r round(medDosage$p.value, 4)`), and the lowest dosage (dose == 0.5; $p=$ `r round(lowDosage$p.value, 4)`).

## Confidence Intervals
95% Confidence Intervals for the difference in mean Odontoblast length are also shown below.
Intervals are shown using Vitamin C as the control and OJ as the treatment.

```{r, echo = FALSE}
confInt <- data.frame(dose = c(0.5, 1, 2),
                      diff = -c(diff(lowDosage$estimate), diff(medDosage$estimate), diff(highDosage$estimate)),
                      lower = c(lowDosage$conf.int[1], medDosage$conf.int[1], highDosage$conf.int[1]),
                      upper = c(lowDosage$conf.int[2], medDosage$conf.int[2], highDosage$conf.int[2]))
ggplot(confInt, aes(x = as.factor(dose), y = diff)) +
  geom_point() +
  geom_errorbar(aes(ymax =upper, ymin=lower)) +
  labs(x = "Dosage", y = "Confidence Interval")
```

# Session Info

```{r}
sessionInfo()
```